
\section{PIFO Queue}

When using a PIFO queue, the scheduling decision is made at the time of enqueue. Packets are pushed into an arbitrary location within the queue, based on their computed rank, and are always dequeued from the head. This abstraction can be decomposed into two components, a programmable rank computation and a fixed scheduler. The programmable rank computation is responsible for producing the packet's priority indicator, or rank. This computation must be performed atomically for each packet and can be expressed in a packet processing language, such as P4~\cite{p4:2014}. The fixed scheduler then uses the resulting rank value to determine the packet's scheduling order relative to all of the packets that are currently in the queue. This paper focuses on exploring the design of the fixed scheduling logic.

For many existing scheduling algorithms, the relative order in which packets are sent does not change with the arrival of new packets. In other words, the scheduling order can be determined at the time the packet is enqueued. This means that the PIFO is a useful abstraction for implementing scheduling policies that have this property. 

PIFOs cannot, however, be used to implement algorithms that require limiting the rate at which packets are dequeued or algorithms that need to arbitrarily change the scheduling order of buffered packets. Despite these constraints, a hierarchical set of PIFOs can be composed into a tree structure to implement hierarchical scheduling algorithms, which can in fact cause reordering of packets that already exist in the queue.
